---
title: "Logistic Regression Demo Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

#### Read Brazilian Beer Data

Recall the beer data contains the liters of beer consumed as a function of precip, maximum temperature, and weekend. Now consider modeling whether the day is a weekend.

```{r messages = F}
library(readr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(arm)
library(rbin)
beer <- read_csv("http://math.montana.edu/ahoegh/Data/Brazil_cerveja.csv")
beer <- beer %>% mutate(weekend_fact = as.factor(weekend))
```


#### Data Viz

It seems likely that consumption patterns would change depending on whether the day was a weekend. 

It does not seem plausible that the weather would be different on the weekends. 

#### Fitting Model

Consider two models: what is the difference between these?
```{r}
glm_beer <- glm(weekend ~ consumed, data = beer)
display(glm_beer)
```



```{r}
glm_beer2 <- glm(weekend_fact ~consumed, data = beer, family = binomial)
display(glm_beer2)
```


### Coefficient interpretations

```{r}
beer <- beer %>% mutate(scaled_consumed = scale(consumed))
glm_beer_scaled <- glm(weekend_fact ~scaled_consumed, data = beer, family = binomial)
beta_mean <- as.numeric(round(exp(coef(glm_beer_scaled)['scaled_consumed'])))
display(glm_beer_scaled)

```

__Intercept:__ The intercept can be interpreted as the the logit scaled probability of success at the zero value of the other parameters.


__scaled_consumed:__ 
All three intepretations are provided, but typically choose one.

_log odds:_ there is evidence that the mean log odds for scaled consumption is positive

_odds:_ there is evidence that the mean odds for scaled consumption is greater than one 

_probability:_ With a one sd increase from mean consumption the probability that the day is a weekend day increases 



### Diagnostics

### Prediction

```{r}
train_ids <- sample(nrow(beer),round(nrow(beer) * .7))
test_ids <- (1:nrow(beer))[!(1:nrow(beer)) %in% train_ids]
train_beer <- beer %>% slice(train_ids)
test_beer <- beer %>% slice(test_ids)

train_glm <- glm(weekend ~ scaled_consumed, family = binomial, data = train_beer)
```
